---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
#Summary
This document was produced as the first project for Coursera's Reproducible Research course.

## Loading and preprocessing the data
The working directory is set and the data is read from the .csv file.
```{r, echo = TRUE}
  setwd("/Users/mistermaxx/Documents/work/personal/Coursera/Repro_Res/")
  
  # read data from file
  activity.file.data <- read.csv("activity.csv")

```


## What is mean total number of steps taken per day?
The aggregate() function is used to summarize the data.
```{r, echo = TRUE, eval = TRUE}
  # calc distrubtion of steps
  steps.data <- aggregate(steps ~ date, activity.file.data, sum)
```

Using ggplot, a histogram of the data is created.
```{r, echo = TRUE, eval = TRUE}
  #create histogram with ggplot
  histogram.steps.data <- ggplot(data = steps.data, aes(steps.data$steps)) 
  histogram.steps.data + geom_histogram(color = "black", fill = "pink", bins = 10, binwidth = 6000) + labs(x = "Steps", y = "Count", title = "Distribution: Total Steps")
  
    # calc, report mean & median
  mean.steps <- mean(steps.data$steps)
  median.steps <- median(steps.data$steps)
  
  print(mean.steps)
  print(median.steps)

```


# What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
